
---
title: "Journey Through `tyecon`: Predictions on Recipe Nutritions"
date: 2022-07-02
categories: [programming, data science]
tags: [R, tyecon]
---

## Introduction

In this second part of the "journey through `tyecon`" series, I want to
showcase the package's facilities in simplifying tasks common to the predictive
modelling part of data analysis. I will be using the *Food.com* dataset as
before.

## Setup and Data Import

```{r setup, message = FALSE}
library(tidyverse)
library(tidyselect)
library(magrittr)
library(vroom)
library(tyecon)
library(rsample)
library(yardstick)
library(glmnet)
library(earth)
library(pls)

knitr::opts_chunk$set(fig.path = "")
knitr::opts_chunk$set(dev = 'svg')
theme_set(theme_light())
set.seed(123)
```

```{r data-import, message = FALSE}
recipes <- vroom("~/Workspace/foodrecipes/RAW_recipes.csv")
interacts <- vroom("~/Workspace/foodrecipes/RAW_interactions.csv")
```


Nutrition values, except for calories, are all percent of daily value, a daily
quota filled by the amount of the percentage by the consumption of the specific
food.

```{r}
nutrs <- recipes %->% {
  ids <- id
  select(nutrition)
  mutate(nutrition = str_replace_all(nutrition, "[\\[\\]\\']", ""))
  separate(nutrition,
    into = c(
      "calories", "sat_fat", "sugar", "sodium", "protein", "tot_fat", "carbs"
    ),
    sep = ", ", convert = TRUE
  )
  # Center and scale too
  mutate(across(everything(), \(col) scale(col)[, 1]))
  set_attr("ids", ids)
}
nutrs
```

And splitting into training and validation:

```r
nutrs_train:nutrs_test %<-%
  (initial_split(nutrs) %>% {
    list(training(.), testing(.))
  })
```

## Predicting Protein from PDV

Of course, since this is an index, it should just work well.

### Linear model

```{r}
lm_model <- lm(protein ~ .,
  data = nutrs_train) %$>% {
  model <- .
  summary(.) %$>% {
    coefs <- as_tibble(coefficients, rownames = "term")
    stats <-
      tibble(r.squared, sigma, df = fstatistic[2], fstat = fstatistic[1])
  }
  pred_metrics <- {
    pred_tibble <- tibble(
      truth = fitted.values(.) + residuals(.), estimate = fitted.values(.)
    )
    map_dfr(c(rmse, rsq, rsq_trad, msd, mae), ~ .(pred_tibble, truth, estimate))
  }
}
lm_model
```

Predictions on the test data:

```{r}
map_dfr(c(rmse, rsq, rsq_trad, msd, mae), ~ .(
  tibble(
    truth = nutrs_test$calories, estimate = predict(lm_model$model, nutrs_test)
  ),
  truth, estimate
))
```

### Multiple models together

We first build the general interface function for our loop over models:

```{r}
lasso <- partial(glmnet, alpha = 1)
ridge <- partial(glmnet, alpha = 0)

# prefix with dot dot to avoid name collision
regression <- convoke(
  list(formula, data),
  lasso(
    x = model.matrix(formula, data)[, -1],
    y = data[[all.vars(formula)[1]]]
  ),
  ridge(
    x = model.matrix(formula, data)[, -1],
    y = data[[all.vars(formula)[1]]]
  ),
  earth(formula, data),
  plsr(formula, data = data)
)
regression
```

And also the predict function:


```{r}
pred_vec <- function(x, newdata, formula, ...) UseMethod("pred_vec")
pred_vec.elnet <- function(x, newdata, formula, ...) {
  predict(x, newx = model.matrix(formula, newdata)[, -1], ...)[, 1]
}
pred_vec.earth <- function(x, newdata, formula, ...) {
  predict(x, newdata = newdata, ...)[, 1]
}
pred_vec.mvr <- function(x, newdata, formula, ...) {
  retval <- predict(x, newdata = newdata, ...)
  retval[, , dim(retval)[[3]]]
}

pred_vec_regs <- conflate(pred_vec(x, newdata, formula))
pred_vec_regs
```

And the main loop:

```{r}
models_preds <- control(
  {
    holdout <- assessment(fold$splits)

    model <- regression(protein ~ .,
      analysis(fold$splits),
      interface = interface, lasso.lambda = lambda,
      earth.degree = degree, plsr.ncomp = 6
    )

    holdout$.fit <- pred_vec_regs(model, holdout, protein ~ .)

    list(model = model, rmse = rmse(holdout, protein, .fit)[[".estimate"]])
  },
  fold = transpose(vfold_cv(nutrs_train, 5)) ~ 1,
  interface = c("lasso", "earth", "plsr") ~ 2,
  lambda = c(0.1, 0.2) ~ 3,
  degree = c(1, 2) ~ 4,
  ncomp = c(5, 6) ~ 5,
  .refiner = ~ confine(.,
    lambda = interface %in% c("lasso", "ridge"), degree = interface == "earth",
    ncomp = interface == "plsr"
  ),
  .unnest_value = TRUE
)
models_preds
```

Now we can choose the appropriate model by testing the data on the validation
set:

```{r}
models_chosen <- group_by(models_preds, across(-c(fold, model, rmse))) %>%
  summarise(model = list(first(model)), rmse = mean(rmse), .groups = "drop")
models_chosen
```

And the validation set:

```{r}
mutate(rowwise(models_chosen), valid_rmse =
  {
    valid_set <- nutrs_test
    valid_set$.fit <- pred_vec_regs(model, valid_set, protein ~ .)
    rmse(valid_set, protein, .fit)[[".estimate"]]
  }
)
```


